---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(dplyr) 
library(ggplot2)
unzip(zipfile = "activity.zip") 
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date)
```
## What is mean total number of steps taken per day?
```{r}
daily_steps <- activity %>% group_by(date) %>%
    summarise(steps_per_day = sum(steps, na.rm = TRUE))

```
1. Histogram: Total number of steps taken each day 
```{r}
hist(daily_steps$steps_per_day, main= "Histogram of Daily Steps", xlab = "Daily Steps", breaks = 20)
```

2. **Mean** and **median** total number of steps taken per day
```{r}
meansteps <- mean(daily_steps$steps_per_day)
```
The mean is `r meansteps`
```{r}
mediansteps <- median(daily_steps$steps_per_day)
```
The median is `r mediansteps`

## What is the average daily activity pattern?
1. Average steps in 5-minute interval    
```{r}
steps_interval <- activity %>%
        group_by(interval) %>%
        summarize(intervalsteps = mean(steps, na.rm = TRUE)) 
plot(steps_interval$interval, steps_interval$intervalsteps, type = "l", xlab = "5-minute Interval", ylab = "Average steps", main = "Average steps by 5-minute intervals")
```
2. Maximum average steps in a 5-minute interval
```{r}
maxsteps <- max(steps_interval$intervalsteps)
```
The maximum average steps in a 5-minute interval are `r maxsteps`

## Imputing missing values
1. Finding the number of observations with missing values 
```{r}
missingvalues <- sum(is.na(activity$steps))
```
There are `r missingvalues' in the data set. 
2. Create main dataset imputing mean to NA 
```{r}
activity1 <- activity
activity1$steps[is.na(activity1$steps)] <- mean(activity1$steps, na.rm=TRUE)
```
3. Total number of steps df on each day after NA = mean
```{r}
daily_steps1 <- activity1 %>% group_by(date) %>%
    summarise(steps_per_day = sum(steps))

```
1. Histogram: Total number of steps taken each day after imputing 
```{r}
hist(daily_steps1$steps_per_day, main= "Histogram of Daily Steps", xlab = "Daily Steps", breaks = 20)
```

2. **Mean** and **median** total number of steps taken per day after imputing
```{r}
meansteps1 <- mean(daily_steps1$steps_per_day)
```
The mean is `r meansteps1`
```{r}
mediansteps1 <- median(daily_steps1$steps_per_day)
```
The median is `r mediansteps1`

## Are there differences in activity patterns between weekdays and weekends?
