---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(dplyr) 
unzip(zipfile = "activity.zip") 
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date)
```
## What is mean total number of steps taken per day?
```{r}
daily_steps <- activity %>% group_by(date) %>%
    summarise(steps_per_day = sum(steps, na.rm = TRUE))

```
1. Make a histogram of the total number of steps taken each day
```{r}
hist(daily_steps$steps_per_day, main= "Histogram of Daily Steps", xlab = "Daily Steps", breaks = 20)
```

2. Calculate and report the **mean** and **median** total number of steps taken per day
```{r}
mean <- mean(daily_steps$steps_per_day)
```
The mean is `r mean`
```{r}
median <- median(daily_steps$steps_per_day)
```
The median is `r median`

## What is the average daily activity pattern?


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
